Phyloseq Preprocessing

### 2022-01-19 Rmarkdown/preprocessing/Phyloseq_Preprocessing_211104.Rmd
Update taxonomy table via MK's suggestions and create an updated PSO named PSO_tmao_NA

@@

## Taxonomy Table
# Load the taxonomy table
taxonTbl <- read_qza(file = "../../data/raw/from_MaryKable/merged-taxonomy-dada2-retrained.qza", tmp = "../../data/raw/from_MaryKable/Q2Rtemp", rm = F)
names(taxonTbl)
taxonTbl$uuid
head(taxonTbl$data) # this reveals that the taxonomy table has classifications to species level (in few cases) (click the right arrow to see). 
# Make taxonomy table human readable
taxTbl<-taxonTbl$data %>% as_tibble() %>% tidyr::separate(Taxon, sep=";", c("Kingdom","Phylum","Class","Order","Family","Genus","Species"))
# View
taxTbl
```

@@

# Save the human readable taxonomy table
write.csv(taxTbl, "../../data/processed/microbiome/phyloseq_inputs/merged-taxonomy-dada2-retrained.csv")
saveRDS(taxTbl, "../../data/processed/microbiome/phyloseq_inputs/merged-taxonomy-dada2-retrained.rds")
```

@@

```{r Make Phyloseq object- specific TMAO variables}
# Load files (metadata, count table, taxonomy table, tree)
# Metadata
metadata_a <- readRDS("../../data/processed/microbiome/phyloseq_inputs/TMAO_specificVariables_mdn_sampledata_211108.rds")
# Count table
ASVr <- readRDS("../../data/processed/microbiome/phyloseq_inputs/merged_table_dada2_no_mito_chloro_highseqreps_0.2rm_subjectID.rds")
# Taxonomy table
taxtable <- readRDS("../../data/processed/microbiome/phyloseq_inputs/merged-taxonomy-dada2-retrained.rds")
# Phylogenetic Tree, rooted
treeR <- read_qza(file = "../../data/raw/from_MaryKable/tmao-tree-files/rooted-tree-dada2_no-mito-chloro_0.2rm_TMAOsamps_mc1.qza",  tmp = "Q2Rtemp", rm = F)

# Create phyloseq object
PSOtmao <- phyloseq(
  otu_table(ASVr, taxa_are_rows = T), 
  phy_tree(treeR$data), 
  tax_table(as.data.frame(taxtable) %>% dplyr::select(-Confidence) %>% column_to_rownames("Feature.ID") %>% as.matrix()), #moving the taxonomy to the way phyloseq wants it
  sample_data(metadata_a)
)
print(PSOtmao)

# Save the phyloseq object
save(PSOtmao, file = "../../data/processed/microbiome/phyloseq_objects/PSOtmao_211108.RData")
```

>>>>

**Updates:**
On 01-19-2022 I updated the chunks, "Original Data from Mary" and "Make Phyloseq object- specific TMAO variables". When MK was reading my manuscript, she noticed that I have some entries in the taxonomy table that say NA and some that say "g__" or "f__". While these may have different meanings, she suggests simply changing all cases to NA. (Another solution, you could manually differentaite "g__" to "g__ASV1" so that the glomming isn't wrong). I added code (from MK) to consistently change the NAs. Then, I created a new phyloseq object, PSO_tmao_NA. These will be added to the project branch named "update-taxonomy-table" in this GitHub project.

@@

## Taxonomy Table
# Load the taxonomy table
taxonTbl <- read_qza(file = "../../data/raw/from_MaryKable/merged-taxonomy-dada2-retrained.qza", tmp = "../../data/raw/from_MaryKable/Q2Rtemp", rm = F)
names(taxonTbl)
taxonTbl$uuid
head(taxonTbl$data) # this reveals that the taxonomy table has classifications to species level (in few cases) (click the right arrow to see). 
# Make taxonomy table human readable (Nazmul's way and the original way I created the phyloseq objects)
#taxTbl<-taxonTbl$data %>% as_tibble() %>% tidyr::separate(Taxon, sep=";", c("Kingdom","Phylum","Class","Order","Family","Genus","Species"))
# View
#taxTbl

# Update on 01-19-2022
# From Mary Kable Jan.14.2022 via email - she suggests we want all unnamed taxa to be NA. For example, an entry that reads "g__" should just say NA. 
tax <- as.data.frame(taxonTbl$data) %>% dplyr::select(-Confidence) %>%
  column_to_rownames("Feature.ID") %>%
  tidyr::separate(Taxon, c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"), sep = ";") %>%
  as.matrix()

# Trim tax to include only SVs in SV table
taxt <- tax[c(rownames(tax)%in%rownames(countData)),]

# Replace all empty levels with NA for the sake of consistancy
# This prevents inaccurate glomming later
taxt[,grep("Species", colnames(taxt))][grep("s__$", taxt[,grep("Species", colnames(taxt))])] <- NA
taxt[,grep("Genus", colnames(taxt))][grep("g__$", taxt[,grep("Genus", colnames(taxt))])] <- NA
taxt[,grep("Family", colnames(taxt))][grep("f__$", taxt[,grep("Family", colnames(taxt))])] <- NA
taxt[,grep("Order", colnames(taxt))][grep("o__$", taxt[,grep("Order", colnames(taxt))])] <- NA
taxt[,grep("Class", colnames(taxt))][grep("c__$", taxt[,grep("Class", colnames(taxt))])] <- NA
taxt[,grep("Phylum", colnames(taxt))][grep("p__$", taxt[,grep("Phylum", colnames(taxt))])] <- NA
```

@@

# Save the human readable taxonomy table
write.csv(taxTbl, "../../data/processed/microbiome/phyloseq_inputs/merged-taxonomy-dada2-retrained.csv")
saveRDS(taxTbl, "../../data/processed/microbiome/phyloseq_inputs/merged-taxonomy-dada2-retrained.rds")

# Save the human readable taxonomy table with consistent NAs per MK's suggestion - 01/19/2022
write.csv(taxt, "../../data/processed/microbiome/phyloseq_inputs/merged-taxonomy-dada2-retrained_consistentNAs.csv")
saveRDS(taxt, "../../data/processed/microbiome/phyloseq_inputs/merged-taxonomy-dada2-retrained_consistentNAs.rds")
```

@@

```{r Make Phyloseq object- specific TMAO variables}
# Load files (metadata, count table, taxonomy table, tree)
# Metadata
metadata_a <- readRDS("../../data/processed/microbiome/phyloseq_inputs/TMAO_specificVariables_mdn_sampledata_211108.rds")
# Count table
ASVr <- readRDS("../../data/processed/microbiome/phyloseq_inputs/merged_table_dada2_no_mito_chloro_highseqreps_0.2rm_subjectID.rds")
# Taxonomy table
#taxtable <- readRDS("../../data/processed/microbiome/phyloseq_inputs/merged-taxonomy-dada2-retrained.rds") 
taxtable <- readRDS("../../data/processed/microbiome/phyloseq_inputs/merged-taxonomy-dada2-retrained_consistentNAs.rds") 
# Phylogenetic Tree, rooted
treeR <- read_qza(file = "../../data/raw/from_MaryKable/tmao-tree-files/rooted-tree-dada2_no-mito-chloro_0.2rm_TMAOsamps_mc1.qza",  tmp = "Q2Rtemp", rm = F)

# Create phyloseq object
PSOtmao_NA <- phyloseq(
  otu_table(ASVr, taxa_are_rows = T), 
  phy_tree(treeR$data), 
  tax_table(as.data.frame(taxtable) %>% 
              #dplyr::select(-Confidence) %>% 
              #column_to_rownames("Feature.ID") %>% 
              as.matrix()), #moving the taxonomy to the way phyloseq wants it
  sample_data(metadata_a)
)
print(PSOtmao_NA)

# Save the phyloseq object
#save(PSOtmao, file = "../../data/processed/microbiome/phyloseq_objects/PSOtmao_211108.RData")
save(PSOtmao_NA, file = "../../data/processed/microbiome/phyloseq_objects/PSOtmao_220119.RData")
```
